New Bayes covariant coupling methods to cope with change of priors in one-vs-one classification

### 2025-08-26 coupling.tex
Fixes from Peter' proofreading

A notable example which uses this principle is the coupling method introduced in \cite{hastie1998classification}. Their method uses Kullback--Leibler divergence 
\begin{align*}
	\divm\nolimits_\textrm{KL} (\boldsymbol{d}, \boldsymbol{d'})= \sum d_i \log (d_i/ d'_i)
\end{align*}

@@

It is useful to consider the graph on $K$ vertices, where vertices correspond to the classes and edges to the selected equations. It cannot have a cycle, otherwise the subsystem of Bradley-Terry equations would be overdetermined for unknowns representing the vertices of the cycle. Therefore it has to be a spanning tree.

@@

A very helpful aid to study Bayes covariant methods is property $BC_{ij}$ (the notation stands for \emph{Bayes covariant when restricted to classes} $i$ and $j$). A coupling method $\boldsymbol{v}$ satisfies $BC_{ij}$ if for any change of priors $\pi$ and any  pairwise data $\boldsymbol{R}$ the odds of classes $C_i$ to $C_j$ predicted by $\boldsymbol{v}(\boldsymbol{R}^\pi)$ and $\bigl(\boldsymbol{v}(\boldsymbol{R})\bigr)^\pi$ are equal. Using earlier notation for odds (see \eqref{eq:alpha}), one requires
\begin{align}
	\alpha_{ij} \bigl(\boldsymbol{v}(\boldsymbol{R}^\pi) \bigr) = \alpha_{ij}\biggl( \bigl(\boldsymbol{v}(\boldsymbol{R}) \bigr)^\pi\biggr)
\end{align}

@@

\begin{prop}
	\label{prop:ensemble}
Suppose the coupling methods $\boldsymbol{v}_1, \ldots, \boldsymbol{v}_M$ are Bayes covariant. If $\sum a_m = 1$, then $\boldsymbol{e} = \bigoplus_m a_m \boldsymbol{v}_m$ is also a Bayes covariant.
\end{prop}

>>>>

A notable example which uses this principle is the coupling method introduced in \cite{hastie1998classification}. Their method uses Kullback--Leibler divergence 
\begin{align*}
	\divm\nolimits_\textrm{KL} (\boldsymbol{d}, \boldsymbol{d'})= \sum d_i \log (d_i/ d'_i).
\end{align*}

@@

It is useful to consider the graph on $K$ vertices, where vertices correspond to the classes and edges to the selected equations. It cannot have a cycle, otherwise the subsystem of Bradley--Terry equations would be overdetermined for unknowns representing the vertices of the cycle. Therefore it has to be a spanning tree.

@@

A very helpful aid to study Bayes covariant methods is property $BC_{ij}$ (the notation stands for \emph{Bayes covariant when restricted to classes} $i$ and $j$). A coupling method $\boldsymbol{v}$ satisfies $BC_{ij}$ if for any change of priors $\pi$ and any  pairwise data $\boldsymbol{R}$ the odds of classes $C_i$ to $C_j$ predicted by $\boldsymbol{v}(\boldsymbol{R}^\pi)$ and $\bigl(\boldsymbol{v}(\boldsymbol{R})\bigr)^\pi$ are equal. Using earlier notation for odds (see \eqref{eq:alpha}), one requires
\begin{align}
	\alpha_{ij} \bigl(\boldsymbol{v}(\boldsymbol{R}^\pi) \bigr) = \alpha_{ij}\biggl( \bigl(\boldsymbol{v}(\boldsymbol{R}) \bigr)^\pi\biggr).
\end{align}

@@

\begin{prop}
	\label{prop:ensemble}
Suppose the coupling methods $\boldsymbol{v}_1, \ldots, \boldsymbol{v}_M$ are Bayes covariant. If $\sum a_m = 1$, then $\boldsymbol{e} = \bigoplus_m a_m \boldsymbol{v}_m$ is also Bayes covariant.
\end{prop}

### 2025-08-26 discussion.tex
Fixes from Peter' proofreading

Let us now turn to recommendation to practitioners on how to handle change of priors in one-vs-one probabilistic models. It is possible to choose any of the following approaches

@@

\begin{table}[!ht]
	\begin{tabular}{cm{2.5cm}|m{1.5cm}ccm{1.5cm}|m{1.5cm}}
	%	&& less competitive & \multicolumn{4}{c|}{competitive method} & oracle \\
		\hline 
		&methods % & Hastie-Tibshirani 
			& Wu--Lin--Weng & radial & normal & arboreal & Hinton's oracle \\
		\hline 
		&source % & \cite{hastie1998classification} 
			& \cite{wu2004probability} & \cite{vsuch2015new} & \cite{vsuch2016bayes} & this work & this work \\
		\multirow{5}{*}{\begin{turn}{90}\makecell{exact}\end{turn}}
		&uniqueness % &  yes 
			& yes & yes & yes & yes & yes \\
		&Bradley--Terry consistency %& yes 
			& yes & yes & yes & yes & yes \\
		&canonicity %& yes 
			& no & yes & yes & no & yes \\
		&symmetry % & yes 
			& yes & yes & yes & no & yes \\
		& Bayes covariant %& no 
			& no & no & yes & yes & yes \\
		\hline
		\multirow{5}{*}{\begin{turn}{90}\makecell{non-exact}\end{turn}}
		&Hinton's minimality %& N/A 
			& \multicolumn{5}{c}{see Section \ref{sec:exp1}} \\
		&robustness %& N/A 
			& \multicolumn{5}{c}{see Section \ref{sec:exp1}} \\
		&lack of parameters %& yes 
			& yes & yes & yes & yes & yes \\
		& amounts to a linear system % & no 
			& yes & yes & yes & yes & yes\\
		& linear system of special form % & N/A 
			& no & yes & yes & yes & yes \\
		& iterative method %& yes 
			& yes & yes & no & not needed & not needed\\
		\hline
	\end{tabular}
	\caption{Summary of properties of coupling methods described in Section \ref{sec:coupling}. The method of Hastie-Tibshirani is omitted from evaluation in Section \ref{sec:exp1}, because earlier evaluation showed it lacks robustness compared to the method of Wu--Lin--Weng \cite{wu2004probability}.}
	\label{tab:summaryCoupling}
\end{table}

@@

Based on our findings we see two immediate pathways for future research. First, one can study arboreal methods in more detail, since these are Bayes covariant and as shown in Section \ref{sec:exp2} they often provide performance matching that of Wu--Lin--Weng's method. Secondly, there is a continuum of Bayes covariant methods which we barely sketched. Their properties, parameter selection,  and applicability to practical tasks need to investigated in greater detail.

>>>>

Let us now turn to recommendation to practitioners on how to handle change of priors in one-vs-one probabilistic models. It is possible to choose any of the following approaches:

@@

\begin{table}[!ht]
	\begin{tabular}{cm{2.5cm}|m{1.5cm}ccm{1.5cm}|m{1.5cm}}
	%	&& less competitive & \multicolumn{4}{c|}{competitive method} & oracle \\
		\hline 
		&methods % & Hastie-Tibshirani 
			& Wu--Lin--Weng & radial & normal & arboreal & Hinton's oracle \\
		\hline 
		&source % & \cite{hastie1998classification} 
			& \cite{wu2004probability} & \cite{vsuch2015new} & \cite{vsuch2016bayes} & this work & this work \\
		\multirow{5}{*}{\begin{turn}{90}\makecell{exact}\end{turn}}
		&uniqueness % &  yes 
			& yes & yes & yes & yes & yes \\
		&Bradley--Terry consistency %& yes 
			& yes & yes & yes & yes & yes \\
		&canonicity %& yes 
			& no & yes & yes & no & yes \\
		&symmetry % & yes 
			& yes & yes & yes & no & yes \\
		& Bayes covariant %& no 
			& no & no & yes & yes & yes \\
		\hline
		\multirow{5}{*}{\begin{turn}{90}\makecell{non-exact}\end{turn}}
		&Hinton's minimality %& N/A 
			& \multicolumn{5}{c}{see Section \ref{sec:exp1}} \\
		&robustness %& N/A 
			& \multicolumn{5}{c}{see Section \ref{sec:exp1}} \\
		&lack of parameters %& yes 
			& yes & yes & yes & yes & yes \\
		& amounts to a linear system % & no 
			& yes & yes & yes & yes & yes\\
		& linear system of special form % & N/A 
			& no & yes & yes & yes & yes \\
		& iterative method %& yes 
			& yes & yes & no & not needed & not needed\\
		\hline
	\end{tabular}
	\caption{Summary of properties of coupling methods described in Section \ref{sec:coupling}. The method of Hastie--Tibshirani is omitted from evaluation in Section \ref{sec:exp1}, because earlier evaluation showed it lacks robustness compared to the method of Wu--Lin--Weng \cite{wu2004probability}.}
	\label{tab:summaryCoupling}
\end{table}

@@

Based on our findings we see two immediate pathways for future research. First, one can study arboreal methods in more detail, since these are Bayes covariant and as shown in Section \ref{sec:exp2} they often provide performance matching that of Wu--Lin--Weng's method. Secondly, there is a continuum of Bayes covariant methods which we barely sketched. Their properties, parameter selection,  and applicability to practical tasks need to be investigated in greater detail.

### 2025-08-26 experiment.tex
Fixes from Peter' proofreading

\begin{itemize}
	\item[(a)] Which coupling methods without parameters show good performance across diverse data sets? Is it the orthodox method of Wu--Lin--Weng? An experiment to answer this question is carried out Section \ref{sec:exp1}. 
	\item[(b)] If the answer to a) is a coupling method which is not Bayes covariant, can we find a parametric Bayes covariant method which  matches its performance? An experiment related to this question is carried out in  Sections \ref{sec:exp2} and \ref{sec:exp3}.
	\item[(c)] If the answer to b) is negative, is it better to apply the change of priors before or after coupling? Additionally, how serious is the impact of the lack of Bayes covariance in practice. This question is investigated in Section \ref{sec:exp4}.
\end{itemize}

@@

\begin{figure}[!ht]
	\includegraphics{exp4-truth.pdf}
	\caption{Difference in accuracies of $\boldsymbol{c}_2$ and $\boldsymbol{c}_1$ for radial and Wu--Lin--Weng's coupling method after change of priors.}
	\label{fig:score}
\end{figure}

>>>>

\begin{itemize}
	\item[(a)] Which coupling methods without parameters show good performance across diverse data sets? Is it the orthodox method of Wu--Lin--Weng? An experiment to answer this question is carried out in Section \ref{sec:exp1}. 
	\item[(b)] If the answer to a) is a coupling method which is not Bayes covariant, can we find a parametric Bayes covariant method which  matches its performance? An experiment related to this question is carried out in  Sections \ref{sec:exp2} and \ref{sec:exp3}.
	\item[(c)] If the answer to b) is negative, is it better to apply the change of priors before or after coupling? Additionally, how serious is the impact of the lack of Bayes covariance in practice. This question is investigated in Section \ref{sec:exp4}.
\end{itemize}

@@

\begin{figure}[!ht]
	\includegraphics{exp4-truth.pdf}
	\caption{Difference in accuracies of $\boldsymbol{c}_2$ and $\boldsymbol{c}_1$ for normal, radial and Wu--Lin--Weng's coupling method after change of priors.}
	\label{fig:score}
\end{figure}
